Accelerated Magnetic Resonance Imaging by Generative Adversarial Neural Networks

### 2018-03-13 paper.tex
fix mistake

%% Group authors per affiliation:
%\author{Ohad Shitrit, Tammy Riklin Raviv}
%\address{$^1$Department of Electrical Engineering, 
%	     $^2$The Zlotowski Center for Neuroscience
%	     Ben-Gurion University of the Negev, Israel}
\IEEEtitleabstractindextext{
\begin{abstract}
A main challenge in Magnetic Resonance Imaging (MRI) is speeding-up scan time. Beyond the improvement of patient experience and the reduction of operational costs, faster scans are essential for time-sensitive imaging, such as fetal, cardio or functional MRI, where temporal resolution is important and target movement is unavoidable, yet must be reduced. Current MRI acquisition methods speed-up scan time at the price of lower spatial resolution and costly hardware.
We introduce a practical, software-only framework, based on deep learning, for accelerating MRI acquisition, while maintaining anatomically meaningful imaging. This is accomplished by MRI sub-sampling, while using a generative adversarial networks to directly estimate the missing k-space samples. The generator-discriminator inter-play enables the introduction of an adversarial cost in addition to fidelity loss used for optimizing the reconstruction. Promising results are obtained for up to six-fold acceleration of diverse brain MRIs, from a large publicly available dataset. Clinical usability of the reconstructed images is assessed by performing tissue segmentation and comparing the results to those obtained by using the original fully-sampled images. Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods, including the widely-used Compressed Sensing.
\end{abstract}

@@

\section{Introduction}
\IEEEPARstart{M}{Magnetic} Resonance Imaging (MRI) is a non-ionizing imaging modality, and is therefore widely used in diagnostic medicine and biomedical research. The physical principles of MRI are based on a strong magnetic field and pulses of radio frequency (RF) electromagnetic radiation. 
Images are produced when hydrogen atoms, which are prevalent in living organisms, emit the absorbed RF energy that is then received by antennas in close proximity to the anatomy being examined. Spatial localization of the detected MRI signals is obtained by varying the magnetic field gradients. The discretized RF output is presented in a Fourier space
(called k-space), where the x-axis refers to the frequency and the y-axis to the phase. An inverse fast Fourier transform (IFFT) of the k-space is then used for generating anatomically meaningful MRI scans. Figure \ref{fig:pair_of_k_space_mri} presents k-space traversal patterns used in conventional imaging. Each row of the k-space is acquired after one RF excitation pulse. The number of rows multiplied by the number of slices (z-axis) determines the total scan time.

@@

A straight forward reduction of the scan time can be obtained by sampling fewer slices, thus reducing the spatial resolution in the z-axis. Spatial distances between adjacent slices of fetal MRI or fMRI, for example, are often as large as 0.5 centimeters. Therefore, a significant portion of the potential input is not covered through imaging. On the other hand, under-sampling in the x-y domain leads to aliasing, as predicted by the Nyquist sampling theorem.

@@

The compressed sensing (CS) technique~\cite{donoho2006compressed} allows efficient acquisition and reconstruction of a signal with fewer samples than the Nyquist-Shannon sampling theorem requires, if the signal has sparse representation in a known transform domain. Using CS for MRI reconstruction by sampling a small subset of the k-space grid had been proposed in~\cite{lustig2007sparse}. The underlying assumption is that the undersampling is random, such that the zero-filled Fourier reconstruction exhibits incoherent artifacts that behave similarly to additive random noise. 
The smooth reconstruction obtained by the CS may result in a loss of noise-like texture that can be anatomically meaningful.

%{\color{red}This, however, makes the algorithm less sensitive to fine details and noise like texture, results in smooth reconstruction with loss of diagnostically information, as we will see in section \ref{experiments_section}.}
%This, however, would require specified pulse programming.
Recently, different machine learning techniques were used for MRI acceleration. 
Ravishankar and Bresler~\cite{ravishankar2011mr} proposed a dictionary-based learning method, to exploit the sparsity of overlapping image patches emphasizing local structure, to reconstruct MRI from highly under-sampled k-space data. In~\cite{caballero2014dictionary}, this concept was extended using spatio-temporal patches for dynamic MRI reconstruction.
Compressed manifold learning based on Laplacian eigenmaps has been applied in~\cite{usman2014compressive} for respiratory motion estimation and in~\cite{bhatia2015fast} for cardiac MRI reconstruction.

Deep neural networks (DNN) allow implicitly learning prior information from large datasets. This has been exploited in recent works for MRI reconstruction from sub-sampled data. A convolutional neural network (CNN) was introduced in~\cite{wang2016accelerating} to learn the mapping between zero-filled and fully-sampled MRIs.
In~\cite{Oktay2016}, MRI acceleration by super-resolution (SR) concepts was introduced. One of the key ideas is a siamese CNN structure that allows receiving multiple inputs acquired from different viewing planes. Improved low-resolution (LR) high-resolution (HR) mapping was obtained by residual architecture.
A multi-scale residual network, also known as U-net~\cite{ronneberger2015u}, was used in~\cite{lee2017deep} to address the CS optimization, formulated as a residual regression problem.
Hayun et el.~\cite{hyun2017deep}, applied U-net to perform anti-aliasing of an MRI generated from sub-sampled k-space. The U-net output image is then enhanced by integrating its k-space domain with the originally acquired raw samples. U-net CNN have been also used in~\cite{sandinodeep} for dynamic MRI reconstruction.
The spatio-temporal characteristics of sequential MRI data was recently exploited using recurrent neural network (RNN) by~\cite{qin2017convolutional}. All these works address the reconstruction problem in the image domain rather than the k-space domain.

The proposed framework utilizes recent advances in deep learning, while addressing MRI reconstruction directly from the k-space. Specifically, we use generative adversarial networks (GAN) \cite{goodfellow2014generative,radford2015unsupervised,pathak2016context}.
GANs are based on the inter-play between two networks: a generator and a discriminator. The generator is capable of learning the distribution over a data-base, and sample realizations of this distribution. 
The discriminator is trained to distinguish between `generated' samples and real ones. This powerful combination has been used for MRI acceleration in~\cite{yu2017deep,mardani2017deep}, and also for generating Computed Tomography (CT)-like images from MRIs~\cite{nie2016medical}.
In contrast, here, the generator is used for reconstruction of the entire k-space grid from under-sampled data. Its loss is a combination of an adversarial loss, calculated on the image domain based on the discriminator output, and a fidelity loss with respect to the fully sampled k-space. Unlike~\cite{hyun2017deep}, the entire process, including the IFFT, is carried out and optimized in an end-to-end manner by the proposed GAN architecture. As the MRI data is complex, both real and imaginary components are incorporated in both the generator and the discriminator.

We compare the proposed method reconstructions to those obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and zero-filling. In addition, for comparison, we tested two other networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss; 2. IM-CNN-L2 - a network that performs on the image-space and is optimized to remove the artifacts caused by under-sampling and zero padding.
The method is applied to the Information eXtraction from Images (IXI) dataset, which is a large publicly available dataset, including 3T and 1.5T scans. Raw data is down-sampled such that only 40\%, 25\%, and 16.6\% of the raw samples of each scan are used.
To assess the clinical usability of the reconstructed images we perform an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity Index (SSIM), we perform tissue segmentation and compare the results to those obtained by using the original fully-sampled images. Segmentation compatibility is measured in terms of Dice scores and Hausdorff distances.
The proposed MRI reconstruction outperforms all other tested methods in all evaluation metrics. Moreover, compared to the widely used CS-MRI, the proposed method is faster by two orders of magnitude. In fact, it can work in real time, and is therefore practical for routine uses.

@@

The rest of the paper is organized as follows: Section~\ref{method_section} provides a brief background on MRI acquisition and introduces the network architecture as well as the training methodology. Section~\ref{experiments_section} presents the experimental results. Conclusions and future directions are given in section~\ref{conclusions_section}.

%--------------------- METHOD --------------------%
\section{Method}\label{method_section}
We aim to address MRI reconstruction from under-sampled k-space using DNN. We now present the proposed methodology and the network architecture.
\subsection{Notations and Problem Definition}
Let $\bm{u}$ denote the desired signal, a 2D MRI, obtained by IFFT of the complex k-space signal $s_{0}$, i.e., 
\begin{equation}
\bm{u}=F^{H}s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
Partial k-space sampling, results in MR images that are subject to artifacts and aliasing. 
Figure~\textbf{\ref{fig:pair_of_k_space_mri}}.d exemplifies artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%).
\vspace{-0.5cm}
\begin{figure}[H]
	\centering
	\setlength\tabcolsep{1.5pt}
	\begin{tabular}{cccc}
		\subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp1/example_mask_kspace_with_arrows}

@@

	} & \subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp2/example_samplingMask4.png}

} & \subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp2/example_zeroPadding.png}
}\tabularnewline
\end{tabular}\caption{\textcolor{black}{\footnotesize{}{}(a) raw k-space: the
	arrows illustrate Cartesian sampling; (b) original; (c) sampling mask; (d) reconstructed}}
\label{fig:pair_of_k_space_mri} 
\end{figure}

@@

%--------------------- Experiments --------------------%
\section{Experiments}\label{experiments_section}

The training data consists of 500 3D brain MRI (T1-weighted) scans of different patients from the IXI dataset.\footnote{\url{http://brain-development.org/ixi-dataset/}} The data was acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each 3D volume we extracted 93 2D saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 3D volumes).
To create k-space images for training, inverse orthonormal
2D FFT was applied to the fully-sampled MR images. 
%Unlike our preliminary work~\cite{shitrit2017accelerated}, which employed a sampling mask along the phase axis alone; here, we perform k-space sampling using a 2D random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space data.
We perform k-space sampling using a 2D random Gaussian mask. Thus, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space.

The generator is composed of $5$ blocks of CONV-BatchNorm-ReLU, with output channels $16,32,64,32,8,$ respectively (see Figure~\ref{fig:architectures}). The last CONV layer is composed of two output channels for real and imaginary components. The discriminator is composed of $4$ blocks of CONV-Pool-BatchNorm-LReLU with output channels $8,16,32,16$ and one fully-connected layer. All CONV layers' kernel size is $3\times3$. All weights were initialized using Xavier \cite{glorot2010understanding}. We used the RMSprop~\cite{tieleman2012lecture} solver with a fixed learning rate of $5\mathrm{e}{-6}$ and set $k_{d}$ to 1.

We compared the proposed method reconstructions to those obtained by using a conventional compressed sensing method-CS-MRI \cite{lustig2007sparse} and zero-filling. In addition, we tested two other networks. To assess the use of an adversarial loss, we trained a k-space generator, with the same architecture as the proposed method, with only an L2 loss (CNN-L2). We also implemented the method proposed in~\cite{hyun2017deep} and constructed a network IM-CNN-L2, which is applied directly to the image space rather than the k-space. Specifically, as in~\cite{hyun2017deep}, we used the U-net architecture~\cite{ronneberger2015u} and applied an FFT on each of its output images to get an estimated k-space. The complete k-space is then recovered by integrating the estimated samples with the sampled raw data according to the sampling mask. The same sampling masks were used for all methods.

To demonstrate the clinical usability of the reconstructed images we performed an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity Index (SSIM)~\cite{wang2004image} metrics, we performed tissue segmentation and compared the results to those obtained by using the original fully-sampled images. The segmentation compatibility is measured in terms of Dice scores and Hausdorff distances.

@@

While PSNR and SSIM are widely used, these measures are not sensitive enough to differences in fine details and complex anatomical structure topologies. For example, a model can provide a good reconstruction in the sense of PSNR, yet with blurry edges and therefore loss of high frequency information. This may significantly affect diagnostic usability. We therefore further compare segmentation quality between the fully-sampled and the reconstruction images. Results of the analysis obtained for our method and all others are presented next.

@@

Brain extraction (skull stripping) is an algorithm that delineates the brain boundary. It is a necessary preliminary step for brain analysis algorithms.
We examine the different reconstruction methods by applying the Brain Extraction Tool (BET) \cite{smith2002fast} to both MR reconstructed and original (fully-sampled) images. Compatibility of the skull stripping results between the original and the generated datasets is measured via the Modified Hausdorff Distance (MHD)~\cite{dubuisson1994modified}. 
%defined as follows.
%
%Let $C_1,C_2\in R^2$ denote the brain boundaries extracted by the BET algorithm, using the original and the reconstructed scans, respectively. The MHD between $C_1$ and $C_2$ is:
%
%\begin{equation}
%\begin{array}{cc}
%MHD(C_1,C_2) = 0.5 \left[{|C_1|^{-1} \sum_{\sigma_{1}\in C_1}^{}d(\sigma_1,C_2) + |C_2|^{-1} \sum_{\sigma_{2}\in C_2}^{}d(\sigma_2,C_1)}\right] \\
%d(\sigma_1,C_2) = \underset{\sigma_{2}\in C_{2}}{\min}||\sigma_1-\sigma_2|| \\
%d(\sigma_2,C_1) = \underset{\sigma_{1}\in C_{1}}{\min}||\sigma_2-\sigma_1||
%\end{array}
%\end{equation}
%
%where $\sigma_1,\sigma_2$ are points on the contours $C_1,C_2$, respectively (Figure~\ref{fig:brain_extraction}.c) and $|C_i|,~i=1,2$, are the respective cardinalities.
MHD values for all tested methods and sampling ratios are presented in~Table~\ref{tbl:MHD}. The MHD scores obtained by our methods are better than all others and are significant (p-values~$<0.0001$) for CS-MRI and zero-filled.

@@

\subsection{Brain Segmentation}
We next tested and compared segmentation quality applying a brain tissue partitioning algorithm (FAST \cite{zhang2001segmentation}) to the reconstructed images. Segmentation results were compared to those obtained for the original fully-sampled scans. We believe that these comparisons are indicative of the diagnostic quality of the reconstructed MRIs as tissue segmentation (WM, GM, and CSF) plays a key role in clinical brain analysis.
%Segmentation compatibility demonstrate the quality of the proposed MRI reconstruction with respect to other methods. We create a reference segmentation by extracting the brain (using BET \cite{smith2002fast}) from the original fully-sampled image and perform a brain segmentation algorithm (FAST \cite{zhang2001segmentation}) on it. Then, we did the same process for all the reconstruction methods and compared the segmentation results.
We used the Dice score~\cite{dice1945measures} to measure the compatibility between the segments obtained for the original ($S_{\mbox{\tiny{REF}}}$) and the reconstructed ($S_{\mbox{\tiny{REC}}}$) images, defined as:
\begin{equation}
\begin{array}{c}
DICE(S_{\mbox{\tiny{REF}}}, S_{\mbox{\tiny{REC}}}) = 2\frac{S_{\mbox{\tiny{REF}}} \cap S_{\mbox{\tiny{REC}}}}{S_{\mbox{\tiny{REF}}} + S_{\mbox{\tiny{REC}}}}
\end{array}
\end{equation}
Dice scores are between zero and one, where one indicates complete overlap between the tested segments. Therefore, higher dice scores imply higher compatibility between reconstructed and original images.
Dice scores for all tested methods and sampling factors are presented in~Table~\ref{tbl:DICE}. As for previous tests, our method achieves the best scores for all brain tissues and all sampling factors. Statistically significant values (p-values~$<0.0005$) were found for all methods using sampling factor of 2.5. In addition to the CS-MRI and zero-filled, the differences are statistically significant with p-values~$<0.0001$ for CNN-L2 using sampling factor 4 and for IM-CNN-L2 using sampling factor 6.

@@

Until just a few years ago, CS techniques were considered the state-of-the-art for MRI reconstruction. Recently, deep learning methods, capable of learning imaging samples distribution from training datasets, were shown to obtain reconstruction with comparable or even higher quality. The strength of the proposed DNN is three-fold, combining direct k-space reconstruction, GAN architecture, and end-to-end optimization.
This allows anatomically meaningful reconstruction with high fidelity to the original scan. Extensive comparative analysis shows that our method outperforms all other tested MRI reconstruction methods (CS and other DNN-based) for various quality metrics, showing more robustness to high acceleration factors.

>>>>

%% Group authors per affiliation:
%\author{Ohad Shitrit, Tammy Riklin Raviv}
%\address{$^1$Department of Electrical Engineering, 
%	     $^2$The Zlotowski Center for Neuroscience
%	     Ben-Gurion University of the Negev, Israel}
\IEEEtitleabstractindextext{
\begin{abstract}
A main challenge in Magnetic Resonance Imaging (MRI) is speeding-up scan time. Beyond the improvement of patient experience and the reduction of operational costs, faster scans are essential for time-sensitive imaging, such as fetal, cardio or functional MRI, where temporal resolution is important and target movement is unavoidable, yet must be reduced. Current MRI acquisition methods speed-up scan time at the price of lower spatial resolution and costly hardware.
We introduce a practical, software-only framework, based on deep learning, for accelerating MRI acquisition, while maintaining anatomically meaningful imaging. This is accomplished by MRI sub-sampling, while using a generative adversarial networks to directly estimate the missing k-space samples. The generator-discriminator inter-play enables the introduction of an adversarial cost in addition to fidelity loss used for optimizing the reconstruction. Promising results are obtained for up to six-fold acceleration of diverse brain MRIs, from a large publicly available dataset. Clinical usability of the reconstructed images is assessed by performing tissue segmentation and comparing the results to those obtained by using the original fully-sampled images. Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods, including the widely-used compressed sensing.
\end{abstract}

@@

\section{Introduction}
\IEEEPARstart{M}{Magnetic} Resonance Imaging (MRI) is a non-ionizing imaging modality, and is therefore widely used in diagnostic medicine and biomedical research. The physical principles of MRI are based on a strong magnetic field and pulses of radio frequency (RF) electromagnetic radiation. 
Images are produced when hydrogen atoms, which are prevalent in living organisms, emit the absorbed RF energy that is then received by antennas in close proximity to the anatomy being examined. Spatial localization of the detected MRI signals is obtained by varying the magnetic field gradients. The discretized RF output is presented in a Fourier space
(called k-space), where the x-axis refers to the frequency and the y-axis to the phase. An inverse fast Fourier transform (IFFT) of the k-space is then applied for generating anatomically meaningful MRI scans. Figure \ref{fig:pair_of_k_space_mri} presents k-space traversal patterns used in conventional imaging. Each row of the k-space is acquired after one RF excitation pulse. The number of rows multiplied by the number of slices (z-axis) determines the total scan time.

@@

A straight forward reduction of the scan time can be obtained by sampling fewer slices, thus reducing the spatial resolution in the z-axis. Spatial distances between adjacent slices of fetal MRI or fMRI, for example, are often as large as 0.5 centimeters. Therefore, a significant portion of the potential input is not acquired through imaging. On the other hand, under-sampling in the x-y domain leads to aliasing, as predicted by the Nyquist sampling theorem.

@@

The compressed sensing (CS) technique~\cite{donoho2006compressed} allows efficient acquisition and reconstruction of a signal with fewer samples than the Nyquist-Shannon sampling theorem limit, when the signal has sparse representation in a known transform domain. Using CS for MRI reconstruction by sampling a small subset of the k-space grid had been proposed in~\cite{lustig2007sparse}. The underlying assumption is that the undersampling is random, such that the zero-filled Fourier reconstruction exhibits incoherent artifacts that behave similarly to additive random noise.
While being considered as a leading technique nowadays ,the CS encourages smooth reconstruction which may result in a loss of noise-like texture that can be anatomically meaningful.

%{\color{red}This, however, makes the algorithm less sensitive to fine details and noise like texture, results in smooth reconstruction with loss of diagnostically information, as we will see in section \ref{experiments_section}.}
%This, however, would require specified pulse programming.
Recently, different machine learning techniques were used for MRI acceleration. 
Ravishankar and Bresler~\cite{ravishankar2011mr} proposed a dictionary-based learning method, to exploit the sparsity of overlapping image patches emphasizing local structure, to reconstruct MRI from under-sampled k-space data. In~\cite{caballero2014dictionary}, this concept was extended using spatio-temporal patches for dynamic MRI reconstruction.
Compressed manifold learning based on Laplacian eigenmaps has been applied in~\cite{usman2014compressive} for respiratory motion estimation and in~\cite{bhatia2015fast} for cardiac MRI reconstruction.

Deep neural networks (DNN) allow implicitly learning prior information from large datasets. This has been exploited in recent works for MRI reconstruction from sub-sampled data. A convolutional neural network (CNN) was introduced in~\cite{wang2016accelerating} to learn the mapping between zero-filled and fully-sampled MRIs.
In~\cite{Oktay2016}, MRI acceleration by super-resolution (SR) concepts was introduced. One of the key ideas is a siamese CNN structure that allows receiving multiple inputs acquired from different viewing planes. Improved low-resolution (LR) high-resolution (HR) mapping was obtained by residual architecture.
A multi-scale residual network, also known as U-net~\cite{ronneberger2015u}, was used in~\cite{lee2017deep} to address the CS optimization, formulated as a residual regression problem.
Hayun et el.~\cite{hyun2017deep}, applied U-net to perform anti-aliasing of an MRI generated from sub-sampled k-space. The U-net output image is then enhanced by integrating its k-space domain with the originally acquired raw samples. U-net CNN have been also used in~\cite{sandinodeep} for dynamic MRI reconstruction.
The spatio-temporal characteristics of sequential MRI data was recently exploited using recurrent neural network (RNN) by~\cite{qin2017convolutional}. All these works address the reconstruction problem in the image domain rather than the k-space domain in which the data is directly acquired.

The proposed framework utilizes recent advances in deep learning, while addressing MRI reconstruction directly from the k-space. Specifically, we use generative adversarial networks (GAN) \cite{goodfellow2014generative,radford2015unsupervised,pathak2016context}.
GANs are based on the inter-play between two networks: a generator and a discriminator. The generator is capable of learning the distribution over a data-base, and sample realizations of this distribution. 
The discriminator is trained to distinguish between `generated' samples and true ones. This powerful combination has been used for MRI acceleration in~\cite{yu2017deep,mardani2017deep}, and also for generating Computed Tomography (CT)-like images from MRIs~\cite{nie2016medical}.
In contrast, here, the generator is used for reconstruction of the entire k-space grid from under-sampled data. Its loss is a combination of an adversarial loss, based on the discriminator output, and a fidelity loss with respect to the fully sampled k-space. Unlike~\cite{hyun2017deep}, the entire process, including the IFFT, is carried out and optimized in an end-to-end manner by the proposed GAN architecture. As the MRI data is complex, both real and imaginary components are fed to both the generator and the discriminator.

We compare the proposed method reconstructions to those obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and zero-filling. In addition, for comparison, we tested two other networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss; 2. IM-CNN-L2 - a network that performs on the image-space and is optimized to remove the artifacts caused by under-sampling and zero padding.
The method is applied to the Information eXtraction from Images (IXI) dataset, which is a large publicly available dataset, including 3T and 1.5T scans. Raw data is down-sampled such that only 40\%, 25\%, and 16.6\% of the raw samples of each scan are used.
To assess the clinical usability of the reconstructed images we perform an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity (SSIM) Index, we perform tissue segmentation and compare the results to those obtained by using the original fully-sampled images. Segmentation compatibility is measured in terms of Dice scores and Hausdorff distances.
The proposed MRI reconstruction outperforms all other tested methods in all evaluation metrics. Moreover, compared to the widely used CS-MRI, the proposed method is faster by two orders of magnitude. In fact, it can work in real-time, and it is, therefore, practical for routine uses.

@@

The rest of the paper is organized as follows: Section~\ref{method_section} provides a brief background on MRI acquisition and introduces the network architecture as well as the training methodology. Section~\ref{experiments_section} presents the experimental results. Conclusions and future directions are given in Section~\ref{conclusions_section}.

%--------------------- METHOD --------------------%
\section{Method}\label{method_section}
We aim to address MRI reconstruction from under-sampled k-space using DNN. We now present the proposed methodology and the network architecture.
\subsection{Notations and Problem Definition}
Let $\bm{u}$ denote the desired signal, a 2D MRI, obtained by IFFT of the complex k-space signal $s_{0}$, i.e., 
\begin{equation}
\bm{u}=F^{H}s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
Partial k-space sampling, results in MR images that are subject to artifacts and aliasing. 
Figure~\ref{fig:pair_of_k_space_mri_example} exemplifies artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%)~Figure~\ref{fig:pair_of_k_space_mri_mask}.
\vspace{-0.5cm}
\begin{figure}[H]
	\centering
	\setlength\tabcolsep{1.5pt}
	\begin{tabular}{cccc}
		\subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp1/example_mask_kspace_with_arrows}

@@

	} & \subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp2/example_samplingMask4.png}\label{fig:pair_of_k_space_mri_mask}

} & \subfloat[\textcolor{black}{\scriptsize{}}]{\centering{}\includegraphics[width=1.8cm,height=1.8cm]{include/grp2/example_zeroPadding.png}\label{fig:pair_of_k_space_mri_example}
}\tabularnewline
\end{tabular}\caption{\textcolor{black}{\footnotesize{}{}(a) raw k-space: the
	arrows illustrate Cartesian sampling; (b) original; (c) sampling mask; (d) reconstructed}}
\label{fig:pair_of_k_space_mri} 
\end{figure}

@@

%--------------------- Experiments --------------------%
\section{Experiments}\label{experiments_section}
\subsubsection*{\bf Data and sampling}
The training data consists of 500 3D brain MRI (T1-weighted) scans of different patients from the IXI dataset.\footnote{\url{http://brain-development.org/ixi-dataset/}} The data was acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each 3D volume we extracted 93 2D saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 3D volumes).
To create k-space images for training, inverse orthonormal
2D FFT was applied to the fully-sampled MR images. 
%Unlike our preliminary work~\cite{shitrit2017accelerated}, which employed a sampling mask along the phase axis alone; here, we perform k-space sampling using a 2D random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space data.
We perform k-space sampling using a 2D random Gaussian mask. Thus, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space.

\subsubsection*{\bf Architecture}
The generator is composed of $5$ blocks of CONV-BatchNorm-ReLU, with output channels $16,32,64,32,8,$ respectively (see Figure~\ref{fig:architectures}). The last CONV layer is composed of two output channels for real and imaginary components. The discriminator is composed of $4$ blocks of CONV-Pool-BatchNorm-LReLU with output channels $8,16,32,16$ and one fully-connected layer. All CONV layers' kernel size is $3\times3$. All weights were initialized using Xavier \cite{glorot2010understanding}. We used the RMSprop~\cite{tieleman2012lecture} solver with a fixed learning rate of $5\mathrm{e}{-6}$ and set $k_{d}$ to 1.

\subsubsection*{\bf Comparisons}
We compared the proposed method reconstructions to those obtained by using a conventional compressed sensing method-CS-MRI \cite{lustig2007sparse} and zero-filling. In addition, we tested two other networks. To assess the use of an adversarial loss, we trained a k-space generator, with the same architecture as the proposed method, with only an L2 loss (CNN-L2). We also implemented the method proposed in~\cite{hyun2017deep} and constructed a network IM-CNN-L2, which is applied directly to the image space rather than the k-space. Specifically, as in~\cite{hyun2017deep}, we used the U-net architecture~\cite{ronneberger2015u} and applied an FFT on each of its output images to get an estimated k-space. The complete k-space is then recovered by integrating the estimated samples with the sampled raw data according to the sampling mask. The same sampling masks were used for all methods.

\subsubsection*{\bf Clinical usability}
To demonstrate the clinical usability of the reconstructed images we performed an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity Index (SSIM)~\cite{wang2004image} metrics, we performed tissue segmentation and compared the results to those obtained by using the original fully-sampled images. The segmentation compatibility is measured in terms of Dice scores and Hausdorff distances.

@@

While PSNR and SSIM are widely used, these measures are not sensitive enough to differences in fine details and complex anatomical structure topologies. For example, a model can provide a good reconstruction in the sense of PSNR, yet with blurry edges and therefore loss of high frequency information. This may significantly affect diagnostic usability. We, therefore, further compare segmentation quality between the fully-sampled and the reconstruction images. Results of the analysis obtained for our method and all others are presented next.

@@

Brain extraction (skull stripping) is an algorithm that delineates the brain boundaries. It is a necessary preliminary step for brain analysis algorithms.
We examine the different reconstruction methods by applying the Brain Extraction Tool (BET) \cite{smith2002fast} to both MR reconstructed and original (fully-sampled) images. Compatibility of the skull stripping results between the original and the generated datasets is measured via the Modified Hausdorff Distance (MHD)~\cite{dubuisson1994modified}. 
%defined as follows.
%
%Let $C_1,C_2\in R^2$ denote the brain boundaries extracted by the BET algorithm, using the original and the reconstructed scans, respectively. The MHD between $C_1$ and $C_2$ is:
%
%\begin{equation}
%\begin{array}{cc}
%MHD(C_1,C_2) = 0.5 \left[{|C_1|^{-1} \sum_{\sigma_{1}\in C_1}^{}d(\sigma_1,C_2) + |C_2|^{-1} \sum_{\sigma_{2}\in C_2}^{}d(\sigma_2,C_1)}\right] \\
%d(\sigma_1,C_2) = \underset{\sigma_{2}\in C_{2}}{\min}||\sigma_1-\sigma_2|| \\
%d(\sigma_2,C_1) = \underset{\sigma_{1}\in C_{1}}{\min}||\sigma_2-\sigma_1||
%\end{array}
%\end{equation}
%
%where $\sigma_1,\sigma_2$ are points on the contours $C_1,C_2$, respectively (Figure~\ref{fig:brain_extraction}.c) and $|C_i|,~i=1,2$, are the respective cardinalities.
MHD values for all tested methods and sampling ratios are presented in~Table~\ref{tbl:MHD}. The MHD scores obtained by our methods are better than all others and the differences are statistically significant (p-values~$<0.0001$) for CS-MRI and zero-filled.

@@

\subsection{Brain Segmentation}
We next tested and compared segmentation quality applying a brain tissue partitioning algorithm (FAST \cite{zhang2001segmentation}) to the reconstructed images. Segmentation results were compared to those obtained for the original fully-sampled scans. We believe that these comparisons are indicative of the diagnostic quality of the reconstructed MRIs as tissue segmentation (WM, GM, and CSF) plays a key role in clinical brain analysis.
%Segmentation compatibility demonstrate the quality of the proposed MRI reconstruction with respect to other methods. We create a reference segmentation by extracting the brain (using BET \cite{smith2002fast}) from the original fully-sampled image and perform a brain segmentation algorithm (FAST \cite{zhang2001segmentation}) on it. Then, we did the same process for all the reconstruction methods and compared the segmentation results.
We used the Dice score~\cite{dice1945measures} to measure the compatibility between the segments obtained for the original ($S_{\mbox{\tiny{REF}}}$) and the reconstructed ($S_{\mbox{\tiny{REC}}}$) images, defined as:
\begin{equation}
\begin{array}{c}
DICE(S_{\mbox{\tiny{REF}}}, S_{\mbox{\tiny{REC}}}) = 2\frac{S_{\mbox{\tiny{REF}}} \cap S_{\mbox{\tiny{REC}}}}{S_{\mbox{\tiny{REF}}} + S_{\mbox{\tiny{REC}}}}
\end{array}
\end{equation}
Dice scores are between zero and one, where one indicates complete overlap between the tested segments. Therefore, higher dice scores imply higher compatibility between reconstructed and original images.
Dice scores for all tested methods and sampling factors are presented in~Table~\ref{tbl:DICE}. As for previous tests, our method achieves the best scores for all brain tissues and all sampling factors. Statistically significant values (p-values~$<0.0005$) were obtained for all methods using sampling factor of 2.5. In addition to the CS-MRI and zero-filled, the differences are statistically significant with p-values~$<0.0001$ for CNN-L2 using sampling factor 4 and for IM-CNN-L2 using sampling factor 6.

@@

Until just a few years ago, CS techniques were considered the state-of-the-art for MRI reconstruction. Recently, deep learning methods, capable of learning imaging samples distribution from training datasets, were shown to obtain reconstruction with comparable or even higher quality. The strength of the proposed DNN is three-fold, combining direct k-space reconstruction, GAN architecture, and end-to-end optimization.
This allows anatomically meaningful reconstruction with high fidelity to the fully-sampled scans. Extensive comparative analysis shows that our method outperforms all other tested MRI methods (CS and other DNN-based) for various quality metrics, showing more robustness to high acceleration factors.
